---
title: "Mapping With more Data - FEMA"
author: "Ricky"
date: "11/10/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
pacman::p_load(
tidyverse,
drat,
maps,
ggplot2,
dplyr,
lubridate,
rgdal,
geojsonio,
tigris,
usmap,
leaflet
)
```

## Load Data
We downloaded FEMA data from this link: <a href="https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1">https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1</a>

This .csv file include detailed public assistance funded projects information. We selected the the Sever Storm(s), Hurricane and Coastal Storm incident-types in 2009-2018 for this analysis.

This dataset is prepared for county-level plot.
```{r echo=FALSE}
hrcc <- read.csv("hrc_county.csv", head = TRUE)
# hrcs <- read.csv("hrc_statewide.csv", head = TRUE)
head(hrcc[,c(1:3,7,9:12)])
```
Here are some explanation from FEMA website:

projectAmount("Project Amount"): The estimated total cost of the Public Assistance grant project in dollars, without administrative costs. This amount is based on the damage survey.

federalShareObligated("Federal Share Obligated"): The Public Assistance grant funding available to the grantee (State) in dollars, for sub-grantee's approved Project Worksheets.

totalObligated("Total Obligated"): The federal share of the Public Assistance grant eligible project amount in dollars, plus grantee (State) and sub-grantee (applicant) administrative costs. The federal share is typically 75% of the total cost of the project.

## Mapping
In this plot, we take all years(2009-2018) data, calculated the summary of obligated amount of each county.
```{r include=FALSE}

hrcc$Fips <- str_pad(hrcc$Fips, 5, side = "left", "0")

# year state obligated
```

```{r}
# total obligated amount in thousands
hrccJ <- hrcc %>% group_by(Fips) %>% 
  summarise(state = unique(state), county = unique(county), 
            TotalAmount = sum(totalObligated)/1000) %>%
  rename(GEO_ID = Fips)
```

In order to give an obvious plot of each county, we delete some outlier points.
```{r include=FALSE}
outValue <- boxplot(hrccJ$TotalAmount)$out
```

```{r echo=TRUE}
# outliers
outliers <- hrccJ %>% filter(TotalAmount %in% outValue)
head(outliers)
hrccJ <- hrccJ %>% filter(!(TotalAmount %in% outValue))
```

```{r include=FALSE}
# json file
loadJ <- geojsonio::geojson_read("gz_2010_us_050_00_5m.json", what = "sp")

# match fips
loadJ$GEO_ID <- loadJ$GEO_ID %>% substr(start = 10, stop = 14)

# join two files
hrccJoin <- geo_join(loadJ, hrccJ, by = "GEO_ID", how = "inner")

# color pal
pal <- colorNumeric("Blues", domain = hrccJoin$TotalAmount)
```

Also add some popups to show the State, County and Total amount(in thousands).
```{r}
# popup 
i_popup <- paste0("<strong>State: </strong>", hrccJoin$state, "<br>", 
                  "<strong>County: </strong>", hrccJoin$county, "<br>", 
                  "<strong>Total Obligated Amount: </strong>", round(hrccJoin$TotalAmount,2) ,
                  " ","thousands") 
```

Plot the leaflet.
```{r}
# leaflet plot
leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addProviderTiles(providers$Stamen.TonerLines,
                   options = providerTileOptions(opacity = 0.75)) %>%
  setView(-89.275673, 37.098, zoom = 4) %>%
  addPolygons(data = hrccJoin,
              fillColor = ~pal(TotalAmount), 
              color = "#BDBDC3",
              fillOpacity  = 0.5, 
              smoothFactor = 0.2,
              weight = 1,
              popup = i_popup) %>%
  addLegend(pal = pal,
            values = hrccJoin$TotalAmount,
            position="bottomright",
            title = paste0("Total Obligated Amount",
                           '<br>',"in thousands (2009-2018)"))
```